---
title: "Result 2"
lof: true
lot: true
format: 
   typst:
     toc: true
     fontsize: 11pt
     section-numbering: 1.1.1.
     columns: 1
     papersize: a4
     # css-property-processing: none
     margin: 
        x: 0.75in
        y: 0.75in
execute: 
  echo: false
  warning: false
  freeze: auto      
# prefer-html: true 
mainfont: CMU Serif
monofont: CMU Typewriter Text
tbl-cap-location: top
keep-typ: true
editor: visual
---

```{r}
#| label: prelims

library(gt)
library(dplyr)
library(data.table)
library(tidyr)
library(purrr)
library(readr)
library(scales)
library(svglite)
library(ggpubr)

source("./utils/theme_set.R")

data_names <- c("Bloom2013", "Bloom2015", "Bloom2019", "Bloom2019_BYxM22", "Bloom2019_RMxYPS163")

plot_preds <- function(pred_df, sample_size=100){
  pred_df %>%
    group_by(Condition, Fold) %>%
    slice_sample(n = sample_size) %>%
    ggplot(., aes(x=Phenotype, y=Preds,)) +
    geom_point(alpha=0.8) +
    stat_smooth(method=lm) +
    stat_cor() +
    facet_wrap(~Condition)
}

fmt_table <- function(matrix){
  matrix %>% 
    as_tibble(rownames="Train dataset") %>% 
    gt(rowname_col="Train dataset") %>% 
    tab_spanner("Test Datasets", everything()) %>% 
    tab_stubhead("Train Datasets") %>% 
    tab_options(
      table.border.bottom.color = "black",
      table.border.bottom.width = px(1.5),
      column_labels.border.top.width = px(2),
      column_labels.border.bottom.width = px(2),
      column_labels.border.top.color = "black",
      column_labels.border.bottom.color = "black",
      row_group.font.weight = "bold",
      stub.font.weight = "bold"
    ) %>% 
    opt_footnote_marks(marks = "letters") %>% 
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_column_labels()
    ) %>% 
    tab_style(
      style = cell_text(align = "center", weight="bold"),
      locations = cells_column_spanners()
    ) %>%
    tab_style(
      style = cell_text(v_align = "middle", weight = "bold"),
      locations = cells_stubhead()
    ) %>% 
    tab_style(
      style = cell_borders(sides="right", color="black", weight = px(1.5)),
      locations = cells_row_groups(everything())
    ) %>% 
    return
}

fmt_conditionwise_table <- function(df){
  df %>% 
    group_by(Condition) %>% 
      summarise(
        Accuracy_mean = mean(accuracy), Accuracy_std = sqrt(var(accuracy)),
        `AUC-ROC_mean`=mean(auc_roc), `AUC-ROC_std`=sqrt(var(auc_roc)),
        F1_mean = mean(f1), F1_std = sqrt(var(f1))
      ) %>% 
    arrange(`AUC-ROC_mean`) %>% 
    gt(rowname_col = "Condition") %>% 
    fmt_number() %>% 
    cols_merge(
      columns = c(Accuracy_mean, Accuracy_std),
      pattern = "{1} ± {2}"
    ) %>% 
    cols_merge(
      columns = c(`AUC-ROC_mean`, `AUC-ROC_std`),
      pattern = "{1} ± {2}"
    ) %>% 
    cols_merge(
      columns = c(F1_mean, F1_std),
      pattern = "{1} ± {2}"
    ) %>% 
    cols_label(
      Accuracy_mean = 'Accuracy',
      `AUC-ROC_mean` = "AUC-ROC",
      F1_mean = 'F1'
    ) %>% 
    tab_stubhead("Conditions") %>% 
    tab_options(
      table.border.bottom.color = "black",
      table.border.bottom.width = px(1.5),
      column_labels.font.weight = "bold",
      column_labels.border.top.width = px(2),
      column_labels.border.bottom.width = px(2),
      column_labels.border.top.color = "black",
      column_labels.border.bottom.color = "black",
      row_group.font.weight = "bold",
      stub.font.weight = "bold"
    ) %>% 
    data_color(
      palette = "viridis"
    ) %>% 
    opt_table_font(
      font = "CMU Serif"
    ) %>% 
    return
}

fmt_conditionwise_table_reg <- function(df){
  df %>% 
    group_by(Condition) %>% 
      summarise(
        `R-Squared` = mean(r2_score), `R-squared_std` = sqrt(var(r2_score)),
        MSE=mean(mean_squared_error), `MSE_std`=sqrt(var(mean_squared_error)),
        Correlation = mean(pearson_correlation), Correlation_std = sqrt(var(pearson_correlation))
      ) %>% 
    arrange(MSE) %>% 
    gt(rowname_col = "Condition") %>% 
    fmt_number() %>% 
    cols_merge(
      columns = c(`R-Squared`, `R-squared_std`),
      pattern = "{1} ± {2}"
    ) %>% 
    cols_merge(
      columns = c(MSE, MSE_std),
      pattern = "{1} ± {2}"
    ) %>% 
    cols_merge(
      columns = c(Correlation, Correlation_std),
      pattern = "{1} ± {2}"
    ) %>% 
    tab_stubhead("Conditions") %>% 
    tab_options(
      table.border.bottom.color = "black",
      table.border.bottom.width = px(1.5),
      column_labels.font.weight = "bold",
      column_labels.border.top.width = px(2),
      column_labels.border.bottom.width = px(2),
      column_labels.border.top.color = "black",
      column_labels.border.bottom.color = "black",
      row_group.font.weight = "bold",
      stub.font.weight = "bold"
    ) %>% 
    data_color(
      palette = "viridis"
    ) %>% 
    opt_table_font(
      font = google_font("CMU Serif")
    ) %>% 
    return
}

```

# Logistic Regression

```{r}
#| label: tbl-logreg-full
#| tbl-cap: Performance of the Logistic Regression Model - AUC-ROC

logreg_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))
logreg_condition <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(logreg_mat)){
  for(colname in colnames(logreg_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_LogReg.csv", sep = "")
    df <- read_csv(
      paste("../data/performances/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    logreg_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    logreg_condition[rowname, colname] <- paste0("../results/result_2/pairwise/","LogReg_", rowname, "_", colname, ".png")
    gtsave(condition.wise.df, paste0("../results/result_2/pairwise/","LogReg_", rowname, "_", colname, ".png"))
  }
}
fmt_table(logreg_mat)
```

```{r}
#| label: tbl-logreg-conditionwise

logreg_condition %>% 
  fmt_table() %>% 
  fmt_image(
    columns = data_names,
    path="./"
  )
```

# Support Vector Machine

```{r}
#| label: tbl-svm
#| tbl-cap: Performance of the Support Vector Machine - AUC-ROC


svm_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))
svm_mat_condition <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(svm_mat)){
  for(colname in colnames(svm_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_SVM.csv", sep = "")
    df <- read_csv(
      paste("../data/performances/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    svm_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/pairwise/","SVM_", rowname, "_", colname, ".png"))
  }
}
fmt_table(svm_mat)
```

# Random Forest

```{r}
#| label: tbl-rf
#| tbl-cap: Performance of the Random Forest Model - AUC-ROC


rf_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(rf_mat)){
  for(colname in colnames(rf_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_RandomForest.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    rf_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/pairwise/","RF_", rowname, "_", colname, ".png"))
  }
}
fmt_table(rf_mat)
```

# Boosting

```{r}
#| label: tbl-boosting
#| tbl-cap: Performance of the GDBT Model - AUC-ROC


boost_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_Boosting.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/pairwise/","Boosting_", rowname, "_", colname, ".png"))
  }
}
fmt_table(boost_mat)
```

## Geno Only

```{r}
#| label: tbl-boosting-geno
#| tbl-cap: Performance of the GDBT Model - AUC-ROC - Genotype only

boost_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_Boosting.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/Geno_only/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/geno_pairwise/","Boosting_", rowname, "_", colname, ".png"))
  }
}
fmt_table(boost_mat)

```

## Chem Only

```{r}
#| label: tbl-boosting-chem
#| tbl-cap: Performance of the GDBT Model - AUC-ROC - Chem only

boost_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_Boosting.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/Chem_only/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/chem_pairwise/","Boosting_", rowname, "_", colname, ".png"))
  }
}
fmt_table(boost_mat)
```

# Baseline - Most frequent

```{r}
#| label: tbl-dummy-mf
#| tbl-cap: Performance of the dummy most frequent Model - AUC-ROC - 

boost_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_most_frequent.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/Dummies/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/dummy/","Most_Frequent_", rowname, "_", colname, ".png"))
  }
}
fmt_table(boost_mat)
```

# Baseline - Stratified

```{r}
#| label: tbl-dummy-stratified
#| tbl-cap: Performance of the dummy stratified Model - AUC-ROC - 

boost_mat <- matrix(nrow=5, ncol=5, dimnames = list(data_names, data_names))

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_stratified.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/Dummies/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/dummy/","stratified_", rowname, "_", colname, ".png"))
  }
}
fmt_table(boost_mat)
```

# Full models

```{r}
#| label: tbl-full
#| tbl-cap: Performance of the full data trained Model - AUC-ROC 


boost_mat <- matrix(nrow=2, ncol=2, dimnames=list(data_names[1:2], data_names[1:2]))

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_Boosting.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/Full/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    # gtsave(condition.wise.df, paste0("../results/result_2/full/","", rowname, "_", colname, ".png"))
    
    svglite(filename = paste0("../results/result_2/full/","Full_", rowname, "_", colname, ".svg"), height=20, width = 20)
    condition.wise.df %>% as_gtable(plot = TRUE)
    dev.off()
  }
}
fmt_table(boost_mat)
```

# Full models - Genotype only

```{r}
#| label: tbl-full-geno_only
#| tbl-cap: Performance of the full data trained Model - AUC-ROC 


boost_mat <- matrix(nrow=2, ncol=2, dimnames=list(data_names[1:2], data_names[1:2]))

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    filename <- paste("metrics_", rowname, "_", colname, "_Boosting.csv", sep = "")
    df <- read_csv(
      paste("../data/performance_boosting/Final/Full/Geno_only/", filename, sep=""),
      col_types = "fidddd"
    )
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    gtsave(condition.wise.df, paste0("../results/result_2/full/","Geno_only_", rowname, "_", colname, ".png"))
  }
}
fmt_table(boost_mat)
```

# Comparing Full and genotype only models

```{r}
#| label: fig-reg-compare-full-geno
#| fig-cap: Performance of the Genotype only data compared to full model

df1 <- read_csv("../data/performance_regression/full/metrics_Bloom2013_Bloom2015_Boosting.csv") %>%
  mutate(run = 'Combined')
df2 <- read_csv("../data/performance_regression/geno_only/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
  mutate(run = "Genotype only")

df <- bind_rows(df1, df2)

df %>% 
  group_by(Condition, run) %>% 
  summarise(r2_score_mean=mean(r2_score), r2_score_std=sqrt(var(r2_score))) %>% 
ggplot(., aes(x=Condition, y = r2_score_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(x=Condition, ymin=r2_score_mean-r2_score_std, ymax=r2_score_mean+r2_score_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Classification - Varying Sigma

```{r}
#| label: fig-clf-sigma0.25
#| fig-cap: Performance of the sigma 0.25


data_names <- c("Bloom2013", "Bloom2015", "Bloom2019_BYxRM", "Bloom2019_BYxM22", "Bloom2019_RMxYPS163")

boost_mat <- matrix(nrow=5, ncol=5, dimnames=list(data_names, data_names))
file_loc <- "../data/performance_classification/classification/0.25_sigma/"
save_dir <- "../results/result_2/classification_varying_sigma/0.25_sigma/"

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    
    print(rowname)
    print(colname)
    
    filename <- paste("metrics_", rowname, "_", colname, "_Boosting.csv", sep = "")
    df <- read_csv(
      paste(file_loc, filename, sep=""),
      col_types = "fidddd"
    )
    df <- rename(df, auc_roc=roc_auc)
    pred_name <- paste("predictions_", rowname, "_",
                       colname, "_Boosting.parquet",
                       sep = "")
    pred_df <- arrow::read_parquet(
      paste(file_loc, pred_name, sep=""),
    )
    
    # Plotting predictions
    p <- plot_preds(pred_df = pred_df)
    save_plot(paste0(save_dir,"Full_", rowname, "_", colname, "_preds.svg"), height=6, width = 9, units='in', scale=1.5)
    
    
    # Adding to table
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    
    svglite(filename = paste0(save_dir,"Full_", rowname, "_", colname, ".svg"), height=20, width = 20)
    condition.wise.df %>% as_gtable(plot = TRUE)
    dev.off()
  }
}
fmt_table(boost_mat)
```

# Classification - 0.5 Sigma

```{r}
#| label: fig-clf-0.5sigma
#| fig-cap: Performance of the 0.5 trained model


data_names <- c("Bloom2013", "Bloom2015", "Bloom2019_BYxRM", "Bloom2019_BYxM22", "Bloom2019_RMxYPS163")

boost_mat <- matrix(nrow=5, ncol=5, dimnames=list(data_names, data_names))
file_loc <- "../data/performance_classification/classification/0.5_sigma/"
save_dir <- "../results/result_2/classification_varying_sigma/0.5_sigma/"

for(rowname in rownames(boost_mat)){
  for(colname in colnames(boost_mat)) {
    
    print(rowname)
    print(colname)
    
    filename <- paste("metrics_", rowname, "_", colname, "_Boosting.csv", sep = "")
    df <- read_csv(
      paste(file_loc, filename, sep=""),
      col_types = "fidddd"
    )
    df <- rename(df, auc_roc=roc_auc)
    pred_name <- paste("predictions_", rowname, "_",
                       colname, "_Boosting.parquet",
                       sep = "")
    pred_df <- arrow::read_parquet(
      paste(file_loc, pred_name, sep=""),
    )
    
    # Plotting predictions
    p <- plot_preds(pred_df = pred_df)
    save_plot(paste0(save_dir,"Full_", rowname, "_", colname, "_preds.svg"), height=6, width = 9, units='in', scale=1.5)
    
    
    # Adding to table
    mean <- mean(df$auc_roc) %>% round(digits = 2)
    std <- sqrt(var(df$auc_roc)) %>% round(digits = 2)
    boost_mat[rowname, colname] <- paste(mean, "±", std)
    
    condition.wise.df <- fmt_conditionwise_table(df)
    
    svglite(filename = paste0(save_dir,"Full_", rowname, "_", colname, ".svg"), height=20, width = 20)
    condition.wise.df %>% as_gtable(plot = TRUE)
    dev.off()
  }
}
fmt_table(boost_mat)
```

# Comparison Classification

```{r}
#| label: fig-reg-compare-full-geno-clf
#| fig-cap: Performance of the Genotype only data compared to full model - classification


df1 <- read_csv("../data/performance_boosting/Final/Full/metrics_Bloom2013_Bloom2015_Boosting.csv") %>%
  mutate(run = 'Combined')
df2 <- read_csv("../data/performance_boosting/Final/Full/Geno_only/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
  mutate(run = "Genotype only")

df <- bind_rows(df1, df2)

df %>% 
  group_by(Condition, run) %>% 
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sqrt(var(auc_roc))
  ) %>% 
  ggplot(., aes(x=Condition, y = auc_roc_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(x=reorder(Condition, auc_roc_mean), ymin=auc_roc_mean-auc_roc_std, ymax=auc_roc_mean+auc_roc_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(
  "../results/result_2/comparison-bloom2013_on_bloom2015.svg",
  width = 9,
  height = 6,
  units = 'in'
)
 # 0.25 sigma
df1 <- read_csv("../data/performance_classification/classification/0.25_sigma/metrics_Bloom2013_Bloom2015_Boosting.csv") %>%
  mutate(run = 'Combined') %>% 
  rename(auc_roc = roc_auc)
# df2 <- read_csv("../data/performance_boosting/Final/Full/Geno_only/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
#   mutate(run = "Genotype only")

df <- bind_rows(df1, df2)

df %>% 
  group_by(Condition, run) %>% 
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sqrt(var(auc_roc))
  ) %>% 
  ggplot(., aes(x=Condition, y = auc_roc_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(x=reorder(Condition, auc_roc_mean), ymin=auc_roc_mean-auc_roc_std, ymax=auc_roc_mean+auc_roc_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(
  "../results/result_2/comparison-bloom2013_on_bloom2015-0.25_sigma.svg",
  width = 9,
  height = 6,
  units = 'in'
)

# 0.5 Sigma

df1 <- read_csv("../data/performance_classification/classification/0.5_sigma/metrics_Bloom2013_Bloom2015_Boosting.csv") %>%
  mutate(run = 'Combined') %>% 
  rename(auc_roc = roc_auc)
# df2 <- read_csv("../data/performance_boosting/Final/Full/Geno_only/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
#   mutate(run = "Genotype only")

df <- bind_rows(df1, df2)

df %>% 
  group_by(Condition, run) %>% 
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sqrt(var(auc_roc))
  ) %>% 
  ggplot(., aes(x=Condition, y = auc_roc_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(x=reorder(Condition, auc_roc_mean), ymin=auc_roc_mean-auc_roc_std, ymax=auc_roc_mean+auc_roc_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(
  "../results/result_2/comparison-bloom2013_on_bloom2015-0.5_sigma.svg",
  width = 9,
  height = 6,
  units = 'in'
)

```

# Comparing Different Sigmas

```{r}
#| label: fig-clf-compare-3sigma
#| fig-cap: Performance of Classification models of varying sigmas


df1 <- read_csv("../data/performance_classification/classification/0.25_sigma/metrics_Bloom2013_Bloom2015_Boosting.csv") %>%
  mutate(run = '0.25 sigma') %>% 
  rename(auc_roc = roc_auc)
df2 <- read_csv("../data/performance_boosting/Final/Full/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
  mutate(run = "1 sigma") 

df3 <- read_csv("../data/performance_classification/classification/0.5_sigma/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
  mutate(run = "0.5 sigma") %>% 
  rename(auc_roc = roc_auc)

df <- bind_rows(df1, df2, df3)

df %>% 
  group_by(Condition, run) %>% 
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sqrt(var(auc_roc))
  ) %>% 
  ggplot(., aes(x=Condition, y = auc_roc_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(x=Condition, ymin=auc_roc_mean-auc_roc_std, ymax=auc_roc_mean+auc_roc_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(
  "../results/result_2/comparison-bloom2013_on_bloom2015-Sigmas.svg",
  width = 9,
  height = 6,
  units = 'in'
)
```

# Comparison Classification 0.25

```{r}
#| label: fig-clf-compare-0.25
#| fig-cap: Performance of the Genotype only data compared to full model - classification 0.25


df1 <- read_csv("../data/performance_classification/classification/0.25_sigma/metrics_Bloom2013_Bloom2015_Boosting.csv") %>%
  mutate(run = 'Combined') %>% 
  rename(auc_roc=roc_auc)
df2 <- read_csv("../data/performance_classification/classification/0.25_sigma_geno_only/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
  mutate(run = "Genotype only") %>% 
  rename(auc_roc=roc_auc)

df <- bind_rows(df1, df2)

df %>% 
  group_by(Condition, run) %>% 
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sqrt(var(auc_roc))
  ) %>% 
  ggplot(., aes(x=Condition, y = auc_roc_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(x=Condition, ymin=auc_roc_mean-auc_roc_std, ymax=auc_roc_mean+auc_roc_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(
  "../results/result_2/classification_varying_sigma/0.25_comparison_Bloom2013_on_Bloom2015.svg",
  width = 9,
  height = 6,
  units = 'in'
)
```

# Comparison Classification 0.5

```{r}
#| label: fig-clf-compare-0.5
#| fig-cap: Performance of the Genotype only data compared to full model - classification 0.25


df1 <- read_csv("../data/performance_classification/classification/0.5_sigma/metrics_Bloom2013_Bloom2015_Boosting.csv") %>%
  mutate(run = 'Combined') %>% 
  rename(auc_roc=roc_auc)
df2 <- read_csv("../data/performance_classification/classification/0.5_sigma_geno_only/metrics_Bloom2013_Bloom2015_Boosting.csv") %>% 
  mutate(run = "Genotype only") %>% 
  rename(auc_roc=roc_auc)


df <- bind_rows(df1, df2)

df %>% 
  group_by(Condition, run) %>% 
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sqrt(var(auc_roc))
  ) %>% 
  ggplot(., aes(x=Condition, y = auc_roc_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(x=Condition, ymin=auc_roc_mean-auc_roc_std, ymax=auc_roc_mean+auc_roc_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(
  "../results/result_2/classification_varying_sigma/0.5_comparison_Bloom2013_on_Bloom2015.svg",
  width = 9,
  height = 6,
  units = 'in'
)
```

```{r}
#| label: fig-clf-compare-0.5-figure-6a
#| fig-cap: Performance of the Genotype only data compared to full model - classification 0.5

df1 <- read_csv("../data/performance_classification/classification/0.5_sigma/metrics_Bloom2015_Bloom2013_Boosting.csv") %>%
  mutate(run = 'Combined') %>% 
  rename(auc_roc=roc_auc)
df2 <- read_csv("../data/performance_classification/classification/0.5_sigma_geno_only/metrics_Bloom2015_Bloom2013_Boosting.csv") %>% 
  mutate(run = "Genotype only") %>% 
  rename(auc_roc=roc_auc)

df3 <- read_csv("../data/performance_classification/classification/0.5_sigma_chem_only/metrics_Bloom2015_Bloom2013_Boosting.csv") %>% 
  mutate(run = "Chem only") %>% 
  rename(auc_roc=roc_auc)

df <- bind_rows(df1, df2, df3)

df %>% 
  filter(Condition %in% c('CaCl2', 'MgSO4', 'sorbitol', 'galactose')) %>% 
  group_by(Condition, run) %>% 
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sqrt(var(auc_roc))
  ) %>% 
  ggplot(., aes(x=Condition, y = auc_roc_mean, fill=run)) +
  geom_bar(stat = 'identity', position = 'dodge', color='black') +
  geom_errorbar(
    aes(x=Condition, ymin=auc_roc_mean-auc_roc_std, ymax=auc_roc_mean+auc_roc_std),
    position='dodge'
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y='Mean AUC-ROC', x='Condition', fill='Run Type')

save_plot(
  "../results/result_2/comparisons/.5_comparison_Bloom2015_on_Bloom2013.svg",
  width = 9,
  height = 6,
  units = 'in')
```

## Condition-wise significance - all train/test combinations

One table per train/test combination. Each run type contributes 5 folds per condition, so the
two-sided Mann-Whitney U cannot go below p = 0.0079 at n = 5 vs 5, and the folds are resamples of
a single dataset rather than independent replicates - the tests describe the stability of the fold
estimates, they are not inferential about new data. Only FDR adjusted values are reported, and
each table is adjusted within itself.

```{r}
#| label: conditionwise-geno-vs-combined-tables
#| output: asis

clf_dir <- "../data/performance_classification/classification/"
clf_datasets <- c("Bloom2013", "Bloom2015", "Bloom2019_BYxRM")

# Display name for the third cross, which is spelled Bloom2019_BYxRM on disk
short_ds <- function(x) sub("Bloom2019_BYxRM", "Bloom2019", x, fixed = TRUE)

runtype_df <- expand_grid(
    tibble(
      run = c("Genotype only", "Combined"),
      run_id = c("geno", "comb"),
      subdir = c("0.5_sigma_geno_only", "0.5_sigma")
    ),
    Train = clf_datasets,
    Test = clf_datasets
  ) %>%
  pmap(function(run, run_id, subdir, Train, Test) {
    read_csv(
      paste0(clf_dir, subdir, "/metrics_", Train, "_", Test, "_Boosting.csv"),
      col_types = "fidddd"
    ) %>%
      rename(auc_roc = roc_auc) %>%
      mutate(run = run, run_id = run_id, Train = Train, Test = Test)
  }) %>%
  list_rbind()

runtype_tests <- runtype_df %>%
  group_by(Train, Test, Condition) %>%
  summarise(
    mwu_p = suppressWarnings(wilcox.test(
      auc_roc[run == "Combined"],
      auc_roc[run == "Genotype only"],
      alternative = "two.sided"
    )$p.value),
    .groups = "drop"
  ) %>%
  # Each pair is its own table, so BH runs within a pair across its conditions
  group_by(Train, Test) %>%
  mutate(mwu_q = p.adjust(mwu_p, "BH")) %>%
  ungroup()

# Prettify condition names without touching the chemical formulae (CoCl2, MnSO4, ...)
pretty_condition <- function(x) {
  x <- gsub("_", " ", as.character(x))
  ifelse(grepl("^[a-z]", x), paste0(toupper(substr(x, 1, 1)), substring(x, 2)), x)
}

runtype_table_df <- runtype_df %>%
  group_by(Train, Test, Condition, run_id) %>%
  summarise(mean = mean(auc_roc), sd = sd(auc_roc), .groups = "drop") %>%
  pivot_wider(names_from = run_id, values_from = c(mean, sd), names_sep = "_") %>%
  left_join(runtype_tests, by = c("Train", "Test", "Condition")) %>%
  mutate(delta = mean_comb - mean_geno, Condition = pretty_condition(Condition)) %>%
  # Same ordering as the condition-wise figure: ascending genotype only AUC-ROC
  arrange(Train, Test, mean_geno)

conditionwise_gt <- function(df) {
  df %>%
    select(Condition, mean_geno, sd_geno, mean_comb, sd_comb, delta, mwu_q) %>%
    gt(rowname_col = "Condition") %>%
    fmt_number(columns = c(mean_geno, sd_geno, mean_comb, sd_comb, delta), decimals = 3) %>%
    fmt_number(columns = mwu_q, decimals = 4) %>%
    cols_merge(columns = c(mean_geno, sd_geno), pattern = "{1} ± {2}") %>%
    cols_merge(columns = c(mean_comb, sd_comb), pattern = "{1} ± {2}") %>%
    cols_label(
      mean_geno = "Genotype only", mean_comb = "Combined",
      delta = "Δ AUC", mwu_q = "q (FDR)"
    ) %>%
    tab_spanner("Mean AUC-ROC", columns = c(mean_geno, mean_comb)) %>%
    tab_stubhead("Conditions") %>%
    cols_width(
      stub() ~ px(150),
      c(mean_geno, mean_comb) ~ px(130),
      delta ~ px(80),
      mwu_q ~ px(90)
    ) %>%
    cols_align(align = "center", columns = -Condition) %>%
    tab_footnote(
      "Mean ± SD over the 5 folds. The Chem only run type is excluded: its AUC-ROC is constant at 0.5.",
      locations = cells_column_spanners(spanners = "Mean AUC-ROC")
    ) %>%
    tab_footnote(
      "Benjamini-Hochberg FDR adjusted p-value of a two-sided Mann-Whitney U over the 5 folds, Combined vs Genotype only, adjusted across the conditions of this table. At n = 5 vs 5 the smallest attainable unadjusted p is 0.0079.",
      locations = cells_column_labels(columns = mwu_q)
    ) %>%
    tab_footnote(
      "Difference in mean AUC-ROC, Combined - Genotype only.",
      locations = cells_column_labels(columns = delta)
    ) %>%
    opt_footnote_marks(marks = "letters") %>%
    tab_options(
      table.border.bottom.color = "black",
      table.border.bottom.width = px(1.5),
      column_labels.font.weight = "bold",
      column_labels.border.top.width = px(2),
      column_labels.border.bottom.width = px(2),
      column_labels.border.top.color = "black",
      column_labels.border.bottom.color = "black",
      row_group.font.weight = "bold",
      stub.font.weight = "bold",
      data_row.padding = px(4)
    ) %>%
    opt_table_font(font = "CMU Serif")
}

runtype_tables <- runtype_table_df %>%
  group_split(Train, Test) %>%
  map(function(d) {
    list(
      gt = conditionwise_gt(d),
      caption = paste0(
        "Condition-wise performance of ", short_ds(d$Train[1]),
        " trained model on ", short_ds(d$Test[1]), " dataset statistical analysis"
      )
    )
  })

# Rendered inline only while knitting, so the chunk can also be run standalone
if (isTRUE(getOption("knitr.in.progress"))) {
  for (t in runtype_tables) {
    cat("\n\n", t$caption, "\n\n", sep = "")
    cat(knitr::knit_print(t$gt))
  }
}
```

## Model types - Kruskal-Wallis over every train/test pair

```{r}
#| label: tbl-clf-modeltype-kw
#| tbl-cap: "Mean AUC-ROC by model type for every train/test combination at 0.5 sigma, with a Kruskal-Wallis test across the four model types. The Chem only run type is excluded, as its AUC-ROC is a constant 0.5."

model_specs <- tibble(
  model = c("LogReg", "SVM", "RandomForest", "Boosting"),
  model_id = c("logreg", "svm", "rf", "boost"),
  subdir = c(rep("0.5_sigma_simple_models", 3), "0.5_sigma"),
  # LogReg/SVM name the third dataset Bloom2019; RandomForest/Boosting use Bloom2019_BYxRM
  ds_2019 = c("Bloom2019", "Bloom2019", "Bloom2019_BYxRM", "Bloom2019_BYxRM")
)
model_datasets <- c("Bloom2013", "Bloom2015", "Bloom2019")

model_df <- expand_grid(model_specs, Train = model_datasets, Test = model_datasets) %>%
  pmap(function(model, model_id, subdir, ds_2019, Train, Test) {
    resolve <- function(ds) if (ds == "Bloom2019") ds_2019 else ds
    d <- read_csv(
      paste0(clf_dir, subdir, "/metrics_", resolve(Train), "_", resolve(Test), "_", model, ".csv"),
      col_types = "fidddd"
    )
    # The AUC column is named roc_auc in some files and auc_roc in others, inconsistently
    # even within one directory - normalise on auc_roc.
    names(d)[names(d) == "roc_auc"] <- "auc_roc"
    stopifnot("auc_roc" %in% names(d))
    d %>% mutate(model = model, model_id = model_id, Train = Train, Test = Test)
  }) %>%
  list_rbind()

# One observation per condition (folds averaged), i.e. the condition-level spread the
# figure's error bars show, rather than pooling condition x fold.
model_condition <- model_df %>%
  group_by(Train, Test, model, model_id, Condition) %>%
  summarise(auc_roc = mean(auc_roc), .groups = "drop")

model_tests <- model_condition %>%
  group_by(Train, Test) %>%
  summarise(kw_p = kruskal.test(x = auc_roc, g = factor(model))$p.value, .groups = "drop") %>%
  mutate(kw_q = p.adjust(kw_p, "BH"))

model_table_df <- model_condition %>%
  group_by(Train, Test, model_id) %>%
  summarise(mean = mean(auc_roc), sd = sd(auc_roc), .groups = "drop") %>%
  pivot_wider(names_from = model_id, values_from = c(mean, sd), names_sep = "_") %>%
  left_join(model_tests, by = c("Train", "Test")) %>%
  arrange(Train, Test) %>%
  mutate(Train = paste("Trained on", Train)) %>%
  select(Train, Test, mean_logreg, sd_logreg, mean_svm, sd_svm,
         mean_rf, sd_rf, mean_boost, sd_boost, kw_q)

model_gt <- model_table_df %>%
  gt(rowname_col = "Test", groupname_col = "Train") %>%
  fmt_number(
    columns = c(mean_logreg, sd_logreg, mean_svm, sd_svm, mean_rf, sd_rf, mean_boost, sd_boost),
    decimals = 3
  ) %>%
  fmt_scientific(columns = kw_q, decimals = 2) %>%
  cols_merge(columns = c(mean_logreg, sd_logreg), pattern = "{1} ± {2}") %>%
  cols_merge(columns = c(mean_svm, sd_svm), pattern = "{1} ± {2}") %>%
  cols_merge(columns = c(mean_rf, sd_rf), pattern = "{1} ± {2}") %>%
  cols_merge(columns = c(mean_boost, sd_boost), pattern = "{1} ± {2}") %>%
  cols_label(
    mean_logreg = "LogReg", mean_svm = "SVM",
    mean_rf = "RF", mean_boost = "Boosting",
    kw_q = "q (FDR)"
  ) %>%
  tab_spanner("Mean AUC-ROC", columns = c(mean_logreg, mean_svm, mean_rf, mean_boost)) %>%
  tab_stubhead("Test dataset") %>%
  # No cols_width here: gt turns fixed widths into absolute p{...pt} columns whose sum
  # overruns \linewidth. Left unset, the LaTeX columns size to their content.
  cols_align(align = "center", columns = -Test) %>%
  tab_footnote(
    "Mean ± SD across the conditions of the test dataset, each condition first averaged over its 5 folds.",
    locations = cells_column_spanners(spanners = "Mean AUC-ROC")
  ) %>%
  tab_footnote(
    "Benjamini-Hochberg FDR adjusted p-value of a Kruskal-Wallis test across the four model types, adjusted across the 9 train/test pairs. One observation per condition, so n per model type is the number of conditions in the test dataset (39 for Bloom2013, 18 for Bloom2015, 31 for Bloom2019).",
    locations = cells_column_labels(columns = kw_q)
  ) %>%
  opt_footnote_marks(marks = "letters") %>%
  tab_options(
    table.border.bottom.color = "black",
    table.border.bottom.width = px(1.5),
    column_labels.font.weight = "bold",
    column_labels.border.top.width = px(2),
    column_labels.border.bottom.width = px(2),
    column_labels.border.top.color = "black",
    column_labels.border.bottom.color = "black",
    row_group.font.weight = "bold",
    stub.font.weight = "bold",
    data_row.padding = px(4),
    # Slightly smaller than the condition-wise tables: this one carries six columns
    table.font.size = px(13)
  ) %>%
  opt_table_font(font = "CMU Serif")

model_gt
```

## Encodings - Kruskal-Wallis over every train/test pair

```{r}
#| label: tbl-clf-encoding-kw
#| tbl-cap: "Mean AUC-ROC for models trained on the three genotype encodings, for every train/test combination at 0.5 sigma, with a Kruskal-Wallis test across the encodings."

enc_dir <- "../data/final/performance/classification_encodings/no_linear_tree/sigma_0.5/"

# Legend order of the encodings figure
encoding_specs <- tibble(
  encoding = c("Max", "Sum", "Count"),
  encoding_id = c("max", "sum", "count")
)

encoding_df <- expand_grid(encoding_specs, Train = clf_datasets, Test = clf_datasets) %>%
  pmap(function(encoding, encoding_id, Train, Test) {
    read_csv(
      paste0(enc_dir, encoding_id, "/metrics_", Train, "_", Test, "_Boosting.csv"),
      col_types = "fidddd"
    ) %>%
      rename(auc_roc = roc_auc) %>%
      mutate(encoding = encoding, encoding_id = encoding_id, Train = Train, Test = Test)
  }) %>%
  list_rbind()

# One observation per condition (folds averaged), as in the model type table
encoding_condition <- encoding_df %>%
  group_by(Train, Test, encoding, encoding_id, Condition) %>%
  summarise(auc_roc = mean(auc_roc), .groups = "drop")

encoding_tests <- encoding_condition %>%
  group_by(Train, Test) %>%
  summarise(kw_p = kruskal.test(x = auc_roc, g = factor(encoding))$p.value, .groups = "drop") %>%
  mutate(kw_q = p.adjust(kw_p, "BH"))

encoding_table_df <- encoding_condition %>%
  group_by(Train, Test, encoding_id) %>%
  summarise(mean = mean(auc_roc), sd = sd(auc_roc), .groups = "drop") %>%
  pivot_wider(names_from = encoding_id, values_from = c(mean, sd), names_sep = "_") %>%
  left_join(encoding_tests, by = c("Train", "Test")) %>%
  arrange(Train, Test) %>%
  mutate(Train = paste("Trained on", short_ds(Train)), Test = short_ds(Test)) %>%
  select(Train, Test, mean_max, sd_max, mean_sum, sd_sum, mean_count, sd_count, kw_q)

encoding_gt <- encoding_table_df %>%
  gt(rowname_col = "Test", groupname_col = "Train") %>%
  fmt_number(
    columns = c(mean_max, sd_max, mean_sum, sd_sum, mean_count, sd_count),
    decimals = 3
  ) %>%
  fmt_number(columns = kw_q, decimals = 3) %>%
  cols_merge(columns = c(mean_max, sd_max), pattern = "{1} ± {2}") %>%
  cols_merge(columns = c(mean_sum, sd_sum), pattern = "{1} ± {2}") %>%
  cols_merge(columns = c(mean_count, sd_count), pattern = "{1} ± {2}") %>%
  cols_label(
    mean_max = "Max", mean_sum = "Sum", mean_count = "Count",
    kw_q = "q (FDR)"
  ) %>%
  tab_spanner("Mean AUC-ROC", columns = c(mean_max, mean_sum, mean_count)) %>%
  tab_stubhead("Test dataset") %>%
  cols_align(align = "center", columns = -Test) %>%
  tab_footnote(
    "Mean ± SD across the conditions of the test dataset, each condition first averaged over its 5 folds.",
    locations = cells_column_spanners(spanners = "Mean AUC-ROC")
  ) %>%
  tab_footnote(
    "Benjamini-Hochberg FDR adjusted p-value of a Kruskal-Wallis test across the three encodings, adjusted across the 9 train/test pairs. One observation per condition, so n per encoding is the number of conditions in the test dataset (39 for Bloom2013, 18 for Bloom2015, 31 for Bloom2019).",
    locations = cells_column_labels(columns = kw_q)
  ) %>%
  opt_footnote_marks(marks = "letters") %>%
  tab_options(
    table.border.bottom.color = "black",
    table.border.bottom.width = px(1.5),
    column_labels.font.weight = "bold",
    column_labels.border.top.width = px(2),
    column_labels.border.bottom.width = px(2),
    column_labels.border.top.color = "black",
    column_labels.border.bottom.color = "black",
    row_group.font.weight = "bold",
    stub.font.weight = "bold",
    data_row.padding = px(4)
  ) %>%
  opt_table_font(font = "CMU Serif")

# All eleven tables as LaTeX in one file. Requires \usepackage{booktabs}, \usepackage{array}
# and, for the condition-wise tables, \usepackage{longtable}; gt writes a literal "±" and "Δ",
# swapped here for math macros so the tables also compile under pdfLaTeX rather than only
# XeLaTeX/LuaLaTeX. gt emits no caption of its own, so a \caption macro is inserted directly
# after the opening environment - plain \caption{} inside a table float, \caption{}\\ inside a
# longtable, where the caption is a row of the table itself.
as_tex <- function(gt_tbl, caption, longtable = FALSE) {
  if (longtable) gt_tbl <- tab_options(gt_tbl, latex.use_longtable = TRUE)
  lines <- strsplit(as.character(as_latex(gt_tbl)), "\n", fixed = TRUE)[[1]]
  opener <- if (longtable) "\\begin{longtable}" else "\\begin{table}"
  at <- grep(opener, lines, fixed = TRUE)[1]
  stopifnot(!is.na(at))
  lines <- append(lines, paste0("\\caption{", caption, "}", if (longtable) "\\\\" else ""), after = at)
  x <- paste(lines, collapse = "\n")
  x <- gsub("±", "$\\pm$", x, fixed = TRUE)
  gsub("Δ", "$\\Delta$", x, fixed = TRUE)
}

writeLines(
  c(
    # One condition-wise table per train/test pair, as longtables so they can break across pages
    map_chr(runtype_tables, function(t) {
      paste0("% ", t$caption, "\n", as_tex(t$gt, t$caption, longtable = TRUE), "\n")
    }),
    "% Model types per train/test pair, Kruskal-Wallis",
    as_tex(
      model_gt,
      paste(
        "Mean AUC-ROC by model type for every train/test combination at 0.5 sigma, with a",
        "Kruskal-Wallis test across the four model types. The Chem only run type is excluded,",
        "as its AUC-ROC is a constant 0.5."
      )
    ),
    "",
    "% Genotype encodings per train/test pair, Kruskal-Wallis",
    as_tex(
      encoding_gt,
      paste(
        "Comparison of performance of models trained on different genotype encodings, for every",
        "train/test combination at 0.5 sigma, with a Kruskal-Wallis test across the three",
        "encodings."
      )
    )
  ),
  "../results/result_2/comparisons/significance_tables-0.5_sigma.tex"
)

encoding_gt
```

## Between encodings

```{r}
#| label: fig-clf-compare-encodings
#| fig-cap: Comparison of performance of models trained on different genotype encodings, across every train/test combination - classification 0.5 sigma

enc_dir <- "../data/final/performance/classification_encodings/no_linear_tree/sigma_0.5/"
encodings <- c("max", "sum", "count")
enc_datasets <- c("Bloom2013", "Bloom2015", "Bloom2019_BYxRM")

enc_df <- expand_grid(
    Encoding = encodings,
    Train = enc_datasets,
    Test = enc_datasets
  ) %>%
  pmap(function(Encoding, Train, Test) {
    read_csv(
      paste0(enc_dir, Encoding, "/metrics_", Train, "_", Test, "_Boosting.csv"),
      col_types = "fidddd"
    ) %>%
      rename(auc_roc = roc_auc) %>%
      mutate(Encoding = Encoding, Train = Train, Test = Test)
  }) %>%
  list_rbind()

enc_summary <- enc_df %>%
  # Average the folds within a condition, then summarise across all conditions
  group_by(Encoding, Train, Test, Condition) %>%
  summarise(auc_roc = mean(auc_roc), .groups = "drop") %>%
  group_by(Encoding, Train, Test) %>%
  summarise(
    auc_roc_mean = mean(auc_roc),
    auc_roc_std = sd(auc_roc),
    .groups = "drop"
  ) %>%
  mutate(
    Encoding = factor(Encoding, levels = encodings, labels = c("Max", "Sum", "Count")),
    Train = factor(
      Train,
      levels = enc_datasets,
      labels = paste("Trained on", c("Bloom2013", "Bloom2015", "Bloom2019"))
    ),
    Test = factor(Test, levels = enc_datasets, labels = c("Bloom2013", "Bloom2015", "Bloom2019"))
  )

ggplot(enc_summary, aes(x = Test, y = auc_roc_mean, fill = Encoding)) +
  geom_col(position = position_dodge(0.9), color = "black") +
  geom_errorbar(
    aes(ymin = auc_roc_mean - auc_roc_std, ymax = auc_roc_mean + auc_roc_std),
    position = position_dodge(0.9),
    width = 0.25
  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  facet_wrap(~Train) +
  scale_fill_manual(values = cbPalette[c(1, 2, 3)]) +
  scale_y_continuous(breaks = seq(0, 1, 0.25)) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = "Test Dataset", y = "Mean AUC-ROC", fill = "Encoding") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

save_plot(
  "../results/result_2/comparisons/encodings-comparison-every_combination-sigma_0.5.svg",
  width = 9,
  height = 4,
  units = 'in'
)
```
